---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggoutline

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/rnabioco/ggoutline/branch/master/graph/badge.svg)](https://codecov.io/gh/rnabioco/ggoutline?branch=master)
<!-- badges: end -->

ggoutline allows users to outline groups of data points using ggplot2.

## Installation

You can install the development version of ggoutline from [GitHub](https://github.com/rnabioco/ggoutline) with:

``` r
# install.packages("devtools")
devtools::install_github("rnabioco/ggoutline")
```

## Why outline points?

A scatter plot is a common way to compare two continuous variables. However, when there are thousands of data points, it can be difficult distinguish between groups based on color alone.

```{r setup}
library(ggoutline)
library(cowplot)
library(purrr)
library(ggplot2)
library(dplyr)

th <- theme_cowplot() +
  theme(
    axis.line    = element_blank(),
    panel.border = element_rect(fill = NA, color = "grey85"),
    axis.ticks   = element_line(color = "grey85")
  )

p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = cluster)) +
  th

p +
  geom_point()
```

geom_outline will add outlines around each group

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = "all"
  )
```

To outline all points together regardless of color or group, set outline_position to 'bottom'

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = "bottom"
  )
```

By passing a group name (or vector of names) to outline_position, a sinlge cluster can be highlighted

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = "c1"
  )
```

By specifying `group` within `aes()`, outlines can be added when coloring with a continuous variable

```{r, fig.width = 8, fig.height = 6}
p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = signal, group = cluster)) +
  scale_color_gradientn(colors = c("white", "red")) +
  th

p +
  geom_outline(outline_size = 2.5)
```

Multiple groups can be highlighted by passed a vector to `outline_position`

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = c("c1", "c2")
  )
```

ggoutline 



