---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  message   = FALSE
)
```

# ggoutline

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/rnabioco/ggoutline/branch/master/graph/badge.svg)](https://codecov.io/gh/rnabioco/ggoutline?branch=master)
<!-- badges: end -->

ggoutline allows users to outline groups of data points using ggplot2.

## Installation

You can install the development version of ggoutline from [GitHub](https://github.com/rnabioco/ggoutline) with:

``` r
# install.packages("devtools")
devtools::install_github("rnabioco/ggoutline")
```

## Why use ggoutline?

A scatter plot is a common way to compare two continuous variables. However, when there are thousands of data points, it can be difficult distinguish between groups based on color alone. With ggoutline groups of points can be highlighted.

```{r setup}
library(ggoutline)
library(cowplot)
library(purrr)
library(ggplot2)
library(dplyr, quietly = TRUE)

th <- theme_cowplot() +
  theme(
    axis.line    = element_blank(),
    panel.border = element_rect(fill = NA, color = "grey85"),
    axis.ticks   = element_line(color = "grey85")
  )

p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = cluster)) +
  th

p +
  geom_point()
```

<br>

geom_outline will add outlines around each group

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = "all"
  )
```

<br>

By passing a group name (or vector of names) to outline_position, a sinlge cluster can be highlighted

```{r, fig.width = 8, fig.height = 6}
p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = "c1"
  )
```

<br>

By specifying `group` within `aes()`, outlines can be added when coloring with a continuous variable

```{r, fig.width = 8, fig.height = 6}
p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = signal, group = cluster)) +
  scale_color_gradientn(colors = c("white", "red")) +
  th

p +
  geom_outline(
    outline_size     = 2.5,
    outline_position = c("c1", "c2")
  )
```



