---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  message   = FALSE
)
```

# ggtrace

<!-- badges: start -->
[![R-CMD-check](https://github.com/rnabioco/ggtrace/workflows/R-CMD-check/badge.svg)](https://github.com/rnabioco/ggtrace/actions)
[![Codecov test coverage](https://codecov.io/gh/rnabioco/ggtrace/branch/master/graph/badge.svg)](https://codecov.io/gh/rnabioco/ggtrace?branch=master)
<!-- badges: end -->

## Overview

A scatter plot is a common way to compare two continuous variables. However, when there are thousands of data points, it can be difficult to distinguish between groups based on color alone. ggtrace provides ggplot2 geoms that highlight groups of data points with an outline for emphasis.

<br>

## Installation

You can install the development version of ggtrace from [GitHub](https://github.com/rnabioco/ggtrace) with:

``` r
# install.packages("devtools")
devtools::install_github("rnabioco/ggtrace")
```

<br>

## Basic Usage

`geom_point_trace` accepts graphical parameters normally passed to `geom_point` to control the appearance of data points and outlines. The fill color for each point can be modified with `fill` and the outline color can be modified with `color`. Additional parameters including `size`, `linetype`, `stroke` and `alpha` are also accepted.

```{r, fig.width = 8, fig.height = 6}
library(ggplot2)
library(ggtrace)

p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = cluster)
) + 
  theme_minimal()

p +
  geom_point_trace(
    color    = "black",
    size     = 1,
    linetype = 1,
    alpha    = 1
  )
```

<br>

## Aesthetics

Like other `ggplot2` geoms, variables can be mapped to aesthetic attributes to modify the outline appearance.

```{r}
ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = cluster)
) +
  theme_minimal() +
  geom_point_trace(
    fill   = "black",
    stroke = 2
  )
```

By specifying `group` within `aes()`, outlines can also be added when coloring with a continuous variable.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = signal, group = cluster)
) +
  theme_minimal()

p +
  scale_fill_gradient(low = "white", high = "red") +
  geom_point_trace(stroke = 0.5)
```

Aesthetics can be further modified using the ggplot2 `scale_*` functions.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = sample)
) +
  theme_minimal()

p +
  geom_point_trace(fill = "white") +
  scale_fill_manual(values = c("red", "#0072B2"))
```

<br>

## Position

The 'position' of the outline can be modified with the `outline_position` parameter. This can be 'all', 'bottom', or a predicate selecting the points to outline. By default 'all' groups are outlined.

To only add a single outline around all points plotted, set `trace_position` to 'bottom'.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = cluster)
) +
  theme_minimal()

p +
  geom_point_trace(
    trace_position = "bottom"
  )
```

A subset of data points can be highlighted by passing a predicate to `outline_position`. This must evaluate to `TRUE` or `FALSE` within the context of the input data.

```{r}
p +
  geom_point_trace(
    trace_position = signal < 0
  )
```

The color of background points can be modified using the `background_color` argument.

```{r}
p +
  geom_point_trace(
    trace_position   = signal < 0,
    background_color = "grey75"
  )
```
