---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  message   = FALSE
)
```

# ggoutline

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/rnabioco/ggoutline/branch/master/graph/badge.svg)](https://codecov.io/gh/rnabioco/ggoutline?branch=master)
<!-- badges: end -->

## Installation

You can install the development version of ggoutline from [GitHub](https://github.com/rnabioco/ggoutline) with:

``` r
# install.packages("devtools")
devtools::install_github("rnabioco/ggoutline")
```

<br>

## Rationale

A scatter plot is a common way to compare two continuous variables. However, when there are thousands of data points, it can be difficult to distinguish between groups based on color alone. This is a first attempt at a package that allows groups of data points to be highlighted using ggplot2. This is under development and likely contains numerous bugs.

geom_point_trace will trace each group plotted

```{r, fig.width = 8, fig.height = 6}
library(ggtrace)
library(tidyverse)
library(cowplot)

th <- theme_cowplot() +
  theme(
    axis.line    = element_blank(),
    panel.border = element_rect(fill = NA, color = "grey85"),
    axis.ticks   = element_line(color = "grey85")
  )

p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = cluster)) +
  th

p +
  geom_point_trace(
    trace_color = "black",
    trace_size  = 1
  )
```

<br>

To add a single outline around all points plotted, set `trace_position` to 'bottom'.

```{r, fig.width = 8, fig.height = 6}
p +
  geom_point_trace(
    trace_color    = "black",
    trace_size     = 1,
    trace_position = "bottom"
  )
```

<br>

A single cluster can be highlighted using standard ggplot2 syntax.

```{r}
p +
  geom_point() +
  geom_point_trace(
    data = ~ filter(.x, cluster == "c1"),
    trace_size = 1
  )
```

<br>

By specifying `group` within `aes()`, outlines can be added when coloring with a continuous variable

```{r, fig.width = 8, fig.height = 6}
p <- clusters %>%
  ggplot(aes(UMAP_1, UMAP_2, color = signal, group = cluster)) +
  scale_color_gradientn(colors = c("white", "red")) +
  th

p +
  geom_point_trace(
    size       = 5,
    trace_size = 1
  )
```
