---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  message   = FALSE
)
```

# ggtrace

<!-- badges: start -->
[![R build
status](https://github.com/rnabioco/ggtrace/workflows/R-CMD-check/badge.svg)](https://github.com/rnabioco/ggtrace/actions)
[![Codecov test coverage](https://codecov.io/gh/rnabioco/ggtrace/branch/master/graph/badge.svg)](https://codecov.io/gh/rnabioco/ggtrace?branch=master)
<!-- badges: end -->

## Installation

You can install the development version of ggtrace from [GitHub](https://github.com/rnabioco/ggtrace) with:

``` r
# install.packages("devtools")
devtools::install_github("rnabioco/ggtrace")
```

<br>

## Rationale

A scatter plot is a common way to compare two continuous variables. However, when there are thousands of data points, it can be difficult to distinguish between groups based on color alone. ggtrace provides ggplot2 geoms that highlight groups of data points with and outline for emphasis .

```{r, fig.width = 8, fig.height = 6}
library(ggplot2)
library(ggtrace)

p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = cluster)
  ) + 
  theme_minimal()

p +
  geom_point_trace(
    trace_color = "black",
    trace_size  = 1
  )
```

To add a single outline around all points plotted, set `trace_position` to 'bottom'.

```{r, fig.width = 8, fig.height = 6}
p +
  geom_point_trace(
    trace_color    = "black",
    trace_size     = 1,
    trace_position = "bottom"
  )
```

A single cluster can be highlighted using standard ggplot2 syntax.

```{r}
p +
  geom_point() +
  geom_point_trace(
    data = ~ subset(.x, cluster == "c1"),
    trace_size = 1
  )
```

By specifying `group` within `aes()`, outlines can be added when coloring with a continuous variable

```{r, fig.width = 8, fig.height = 6}
ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = signal, group = cluster)
  ) +
  theme_minimal() +
  scale_color_gradientn(
    colors = c("white", "red")
    ) +
  geom_point_trace(
    size       = 5,
    trace_size = 1
  )
```
