---
title: "geom_point_trace"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  rmarkdown::html_vignette:
    toc_float: true
    toc_depth: 2
    vignette: >
      %\VignetteIndexEntry{V(D)J analysis}
      %\VignetteEncoding{UTF-8}
      %\VignetteEngine{knitr::rmarkdown}
---

## Basic Usage

`geom_path_trace()`, `geom_line_trace()`, and `geom_step_trace()` accept arguments normally passed to `geom_path()`, `geom_line()`, and `geom_step()` with some minor modifications. `fill` controls the line color, `color` controls the outline color, and `stroke` controls the outline width, similar to how filled shapes are modified for other `ggplot2` geoms. Additional parameters including `size`, `linetype` and `alpha` are also accepted and behave normally.

```{r}
library(ggplot2)
library(ggtrace)

dat <- EuStockMarkets %>%
  as_tibble() %>%
  mutate(day = row_number()) %>%
  pivot_longer(-day)

p <- ggplot(dat, aes(day, value, fill = name)) +
  theme_minimal()

p +
  geom_line_trace(
    color    = "black",
    size     = 1,
    linetype = 1,
    alpha    = 1
  )
```

<br>

## Aesthetics

Like other `ggplot2` geoms, variables can be mapped to aesthetic attributes to modify the outline appearance.

```{r}
p <- ggplot(dat, aes(day, value, color = name)) +
  theme_minimal()

p +
  geom_line_trace()
```

<br>

Aesthetics can be further modified using the ggplot2 `scale_*` functions.

```{r}
p +
  geom_line_trace() +
  scale_color_manual(values = c("#E69F00", "#0072B2", "#56B4E9", "#009E73"))
```

<br>

## Position

The 'position' of the outline can be modified with the `outline_position` parameter. This can be 'all' or a predicate selecting the data to outline. By default 'all' groups are outlined.

A subset of data points can be highlighted by passing a predicate to `outline_position`. This must evaluate to `TRUE` or `FALSE` within the context of the input data.

```{r}
p <- ggplot(dat, aes(day, value, color = name, group = name)) +
  theme_minimal()

p +
  geom_line_trace(
    trace_position   = day > 1000,
    size             = 0.5,
    stroke           = 1,
    fill             = "black",
    background_color = "grey75"
  ) +
  scale_color_manual(values = c("#FFC273", "#69A9EA", "#94D3FF", "#5ACCA0"))
```

<br>

```{r}
p +
  geom_line_trace(
    trace_position   = name == "FTSE",
    size             = 0.5,
    stroke           = 1,
    fill             = "grey75",
    background_color = "grey75"
  )
```




