---
title: "geom_point_trace"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{geom_point_trace}
      %\VignetteEncoding{UTF-8}
      %\VignetteEngine{knitr::rmarkdown}
---

## Basic Usage

`geom_point_trace` accepts graphical parameters normally passed to `geom_point` to control the appearance of data points and outlines. The fill color for each point can be modified with `fill` and the outline color can be modified with `color`. Additional parameters including `size`, `linetype`, `stroke` and `alpha` are also accepted.

```{r}
library(ggplot2)
library(ggtrace)

p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = cluster)
) + 
  theme_minimal()

p +
  geom_point_trace(
    color    = "black",
    size     = 1,
    linetype = 1,
    alpha    = 1
  )
```

<br>

## Aesthetics

Like other `ggplot2` geoms, variables can be mapped to aesthetic attributes to modify the outline appearance.

```{r}
ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = cluster)
) +
  theme_minimal() +
  geom_point_trace(
    fill   = "black",
    stroke = 2
  )
```

<br>

By specifying `group` within `aes()`, outlines can also be added when coloring with a continuous variable.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = signal, group = cluster)
) +
  theme_minimal()

p +
  scale_fill_gradient(low = "white", high = "red") +
  geom_point_trace(stroke = 0.5)
```

<br>

Aesthetics can be further modified using the ggplot2 `scale_*` functions.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, color = sample)
) +
  theme_minimal()

p +
  geom_point_trace() +
  scale_color_manual(values = c("red", "#0072B2"))
```

<br>

## Position

The 'position' of the outline can be modified with the `outline_position` parameter. This can be 'all', 'bottom', or a predicate selecting the points to outline. By default 'all' groups are outlined.

To only add a single outline around all points plotted, set `trace_position` to 'bottom'.

```{r}
p <- ggplot(
  clusters,
  aes(UMAP_1, UMAP_2, fill = cluster)
) +
  theme_minimal()

p +
  geom_point_trace(
    trace_position = "bottom"
  )
```

<br>

A subset of data points can be highlighted by passing a predicate to `outline_position`. This must evaluate to `TRUE` or `FALSE` within the context of the input data.

```{r}
p +
  geom_point_trace(
    trace_position = signal < 0
  )
```

<br>

The color of background points can be modified using the `background_color` argument.

```{r}
p +
  geom_point_trace(
    trace_position   = signal < 0,
    background_color = "grey85"
  )
```

<br>

To leave out the outline.

```{r}
p +
  geom_point_trace(
    trace_position   = signal < 0,
    background_color = "grey85",
    color            = NA
  )
```




