---
title: "geom_line_trace"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  rmarkdown::html_vignette:
    toc_float: true
    toc_depth: 2
    vignette: >
      %\VignetteIndexEntry{V(D)J analysis}
      %\VignetteEncoding{UTF-8}
      %\VignetteEngine{knitr::rmarkdown}
---

## Basic Usage

`geom_path_trace()`, `geom_line_trace()`, and `geom_step_trace()` accept arguments normally passed to `geom_path()`, `geom_line()`, and `geom_step()` with some minor modifications. `fill` controls the line color, `color` controls the outline color, and `stroke` controls the outline width, similar to how filled shapes are modified for other `ggplot2` geoms. Additional parameters including `size`, `alpha`,  `linetype`, `linejoin`, `lineend` and `linemitre` are also accepted and behave normally.

```{r}
library(ggplot2)
library(ggtrace)
library(tidyr)

dat     <- as.data.frame(EuStockMarkets)
dat$day <- as.numeric(rownames(dat))
dat     <- pivot_longer(dat, -day)

p <- ggplot(dat, aes(day, value, fill = name)) +
  theme_minimal()

p +
  geom_line_trace(
    color    = "black",
    size     = 1,
    stroke   = 0.5,
    linetype = 1,
    alpha    = 1
  )
```

<br>

## Aesthetics

Like other `ggplot2` geoms, variables can be mapped to aesthetic attributes to modify the outline appearance.

```{r}
p <- ggplot(dat, aes(day, value, color = name)) +
  theme_minimal()

p +
  geom_line_trace()
```

<br>

Aesthetics can be further modified using the ggplot2 `scale_*` functions.

```{r}
clrs <- c(
  CAC  = "#E69F00",
  DAX  = "#0072B2",
  SMI  = "#56B4E9",
  FTSE = "#009E73"
)

p +
  geom_line_trace() +
  scale_color_manual(values = clrs)
```

<br>

## Position

The 'position' of the outline can be modified with the `trace_position` parameter. This can be 'all' or a predicate selecting the data to outline. By default 'all' groups are outlined.

If a predicate is passed to `trace_position`, it must evaluate to `TRUE` or `FALSE` within the context of the input data.

```{r}
p <- ggplot(dat, aes(day, value, color = name)) +
  theme_minimal()

p +
  geom_line_trace(
    trace_position = day > 1500,
    size           = 0.5,
    stroke         = 1,
    fill           = "black"
  )
```

<br>

The color of background lines can be modified using the `background_color` argument.

```{r}
p +
  geom_line_trace(
    trace_position   = day > 1500,
    size             = 0.5,
    stroke           = 1,
    fill             = "black",
    background_color = "grey75"
  )
```

<br>

Here we only highlight names where the second letter is 'A'.

```{r}
p +
  geom_line_trace(
    trace_position   = grepl("^[A-Z]A", name),
    fill             = "black",
    background_color = "grey75"
  )
```

<br>

Outlines can be removed by setting `color` to `NA`.

```{r}
p <- ggplot(dat, aes(day, value, fill = name)) +
  theme_minimal()

p +
  geom_line_trace(
    trace_position   = grepl("^[A-Z]A", name),
    color            = NA,
    background_color = "grey75"
  )
```





